Abstractions for Multi-Sorted Substitutions

### 2023-07-11 paper/paper.tex
Update.

\title{Kit Theory}

@@

  \section{Syntax}
  The following shows a typical intrinsically-scoped syntax of System F:
  \ExampleCode\FUnsortedSyntax

  \texttt{Type}s are indexed by the number of free type variables \texttt{n}.
  \texttt{Expr}essions are additionally indexed by the number of free expression variables \texttt{m}.
  Variables are represented as DeBruijn-indices, where \texttt{Fin n} is the type of \texttt{n} elements.

  This style of syntax comes with multiple drawbacks:
  \begin{itemize}
  \item Bad scaling for automation
  \item Multiple substitions require interaction lemmas
  \end{itemize}

  % Instead we define a single term type which is indexed by the sort of
  % syntax, i.e.\ whether the term is a kind, type, or expression:
  To avoid these drawbacks, we define a single term type \texttt{S~⊢~s} that is indexed by its sort \texttt{s},
  i.e.\ whether the term is a kind, type, or expression, and the sorts of its free variables \texttt{S}:

@@

  \LibCode\KTerms
  \AppCode\FTerms

@@

  A renaming \texttt{S₁ →ᵣ S₂} maps variables from \texttt{S₁} to
  variables from \texttt{S₂}.
  A substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
  terms in \texttt{S₂}.

@@

  \subsection{Unifying Renamings \& Substitutions}
  A well-known problem with DeBruijn-indices is that full substitution
  cannot be defined directly by structural recursion on the term.
  The typical workaround is to define renamings and substitutions
  separately and use renamings to weaken the terms in substitutions
  when going under binders.

@@

  To avoid this duplication,
  McBride\cite{DBLP:journals/jar/BentonHKM12,
  unpublished:mcbride2005kits} introduced \emph{Kits} - an abstraction
  which allows to talk about both renamings and substitutions
  simultaneously.

  The core idea behind kits is that renamings and substitutions only
  differ in what they return:
  \ExampleCode\FExampleSubRen

@@

  \subsection{Library}
  \LibCode\KMap
  \LibCode\KAp
  \LibCode\KExt
  \LibCode\KLift
  \LibCode\KId
  \LibCode\KSingle
  \LibCode\KWeaken
  \LibCode\KEq
  \LibCode\KFunExt
  \LibCode\KIdLift
  % \LibCode\KIdLiftProof
  \LibCode\KKitNotation
  \LibCode\KTraversal
  \LibCode\KKitInstances
  \LibCode\KKitOpen
  \subsection{System F}
  \AppCode\FTraversalOp
  \AppCode\FTraversalId
  \AppCode\FTraversalIdProofInteresting
  % \LibCode\FTraversalIdProofRest
  \AppCode\FTraversal

>>>>

\title{Abstractions for Multi-Sorted Substitutions}

@@

  \section{Syntax}
  \subsection{Single-Sorted Syntax}
  The following shows a typical intrinsically-scoped syntax of System F:
  \ExampleCode\FUnsortedSyntax

  \texttt{Type}s are indexed by the number of free type variables \texttt{n}.
  \texttt{Expr}essions are additionally indexed by the number of free expression variables \texttt{m}.
  Variables \texttt{`\_} are represented as DeBruijn-indices, where \texttt{Fin n} is the type of \texttt{n} elements.

  We identify two drawbacks with this style of syntax:
  \begin{enumerate}
  \item the syntactic categories (\texttt{Kind}, \texttt{Type}, and \texttt{Expr})
    have different types, which makes it difficult to treat them uniformly; and
  \item the different sorts of variables are modeled separately, which requires to define not just
    type-in-type and expression-in-expression substitution, but also type-in-expression substitution.
    Consequently, interaction lemmas between those substitutions are required.
    In general, this leads to a combinatory explosion of
    substitutions, e.g.\ adding kind variables would lead to a total of 6
    different substitutions.
  \end{enumerate}

  \subsection{Multi-Sorted Syntax}
  % Instead we define a single term type which is indexed by the sort of
  % syntax, i.e.\ whether the term is a kind, type, or expression:
  To avoid these drawbacks, we define a single term type \texttt{S~⊢~s} that is indexed by its sort \texttt{s},
  i.e.\ whether the term is a kind, type, or expression, and the sorts of its free variables \texttt{S}:

@@

  \subsection{A Structure for Multi-Sorted Syntax}
  \LibCode\KTerms
  \AppCode\FTerms

@@

  A renaming \texttt{S₁ →ᵣ S₂} maps variables from \texttt{S₁} to
  variables from \texttt{S₂}.
  A substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
  terms containing variables from \texttt{S₂}.

@@

  \subsection{Unifying Renamings \& Substitutions}
  A well-known problem with DeBruijn-indices is that full substitution
  cannot be defined directly by structural recursion on the term.
  The usual workaround is to define renamings and substitutions
  separately and use renamings to weaken the terms in substitutions
  when going under binders.

@@

  To avoid this duplication,
  McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
  introduced an abstraction called \emph{Kits}, which allows to talk about
  both renamings and substitutions simultaneously.

  The core observation leading to this abstraction is that renamings
  and substitutions only differ in their return type:
  \ExampleCode\FExampleSubRen

@@

  \subsection{Library}
  \LibCode\KAp
  \LibCode\KExt
  \LibCode\KLift
  \LibCode\KId
  \LibCode\KSingle
  \LibCode\KWeaken
  \LibCode\KEq
  \LibCode\KFunExt
  \LibCode\KIdLift
  % \LibCode\KIdLiftProof
  \LibCode\KKitNotation
  \LibCode\KTraversal
  \LibCode\KKitInstances
  \LibCode\KKitOpen
  \subsection{System F}
  \AppCode\FTraversalOp
  \AppCode\FTraversalId
  \AppCode\FTraversalIdProofInteresting
  % \LibCode\FTraversalIdProofRest
  \AppCode\FTraversal
